---
title: "Test Results"
format: html
editor: visual
---

#### Load Libraries

```{r, message=FALSE}
library(tidyverse)
library(janitor)
library(vembedr)
```

#### The Story

You test positive for a rare disease that only effects 0.001 (One in one thousand people).

So you ask the doctor:

-   How certain is it that I have this disease?

    -   The test correctly identifies 99% of people that have the disease and only incorrectly identifies 1% of people that don't have the disease

What are the chances that you actually have this disease?

-   Some would say 99%, the accuracy of the test

    -   What does bayes say?

$$
P(B \mid A) = \frac{P(B) L(B \mid A)}{P(A)} 
$$

B \<- Has Disease

```{r}
patients%>% 
  tabyl(has_disease)

```

```{r}
prob_has_disease <- 0.001
```

A \<- Positive test result

```{r}
patients%>% 
  tabyl(test_result)

```

```{r}
prob_has_postest <- 0.0112
```

P(B\|A) - The probability of having the disease given a positive test result

```{r}
patients %>% 
  tabyl(has_disease, test_result) %>% 
  adorn_percentages("row")
```

```{r}
likelihood <- 1.00000000
```

```{r}
(prob_has_disease * likelihood) / prob_has_postest
```

```{r}
0.08928571 * 100 
## about a 9% chance?
```

#### Simulate the Data

```{r}

set.seed(70)  # For reproducibility

# Parameters
n_patients <- 10000  # Total population size
n_diseased <- 10     # Number of patients with the disease
sensitivity <- 0.99  # True positive rate (sensitivity)
false_positive_rate <- 0.01  # False positive rate

# Step 1: Create the DataFrame with patients
patients <- data.frame(
  patient_id = 1:n_patients,
  has_disease = c(rep(1, n_diseased), rep(0, n_patients - n_diseased))  # 10 with the disease, rest without
)

# Shuffle the DataFrame to randomize patient order
patients <- patients %>%
  sample_frac(size = 1)

# Step 2: Simulate the test results based on disease status
patients <- patients %>%
  mutate(
    # Test result is positive if the person has the disease and the test is sensitive,
    # or if they don't have the disease but it's a false positive
    test_result = case_when(
      has_disease == 1 & rbinom(n_patients, size = 1, prob = sensitivity) == 1 ~ "positive",
      has_disease == 0 & rbinom(n_patients, size = 1, prob = false_positive_rate) == 1 ~ "positive",
      TRUE ~ "negative"
    )
  )
```

#### Apply Bayes Theorem in Class

#### Video:

```{r}
embed_url("https://www.youtube.com/watch?v=R13BD8qKeTg")
```

#### DS400 Assignment:

$$
P(\text{have disease} \mid \text{positive second test}) = \frac{P(\text{have disease after first positive}) \cdot P(\text{positive second test} \mid \text{have disease})}{P(\text{positive second test})}
$$

```{r}
## [future code would go here !!]
```

In class, we solved for the following P(B\|A).

$$
P(\text{have disease} \mid \text{positive test}) = \frac{P(\text{has disease}) \cdot P(\text{has disease | positive test result)}}{P(\text{positive test result})}
$$

In the example above, we were solving for the probability of having the disease given a positive test result, which ultimately gave us an answer of 9%. In the context of solving for the probability that a patient has a disease given a second positive test result, I would assume that the probability of the evnt (or the % of it) would increase. This is because a second positive test would provide additional evidence that the patient may actually have the disease. When it comes to using Bayes' theorem again, my best guess would be that the prior probability would have to be updated to account for the likelihood of a second positive result given that the patient has the disease. In a way... would this question be considered to be an application of Bayes Theorem within a Bayes theorem?
