---
title: "NHANES Review & Example"
format: html
editor: visual
---

### Load Libraries

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(car)
library(tidyverse)
library(haven)
library(here)
library(ggplot2)
library(janitor)
```

### Read-in Data:

-   DR1TOT_J.xpt as 'diet_behavior'
-   2017-2018_Hg-Blood.XPT as 'blood_hg'
-   2017-2018_Hg-Urine.XPT as 'urine_hg'
-   2017-2018_Diabetes.XPT as 'diabetes'

```{r}
diet_behavior <- read_xpt(here("data/nhanes/DR1TOT_J.XPT"))
blood_hg <- read_xpt(here("data/nhanes/2017-2018_Hg-Blood.XPT"))
urine_hg <- read_xpt(here("data/nhanes/2017-2018_Hg-Urine.XPT"))
diabetes <- read_xpt(here("data/nhanes/2017-2018_Diabetes.XPT"))
demographics <- read_xpt(here("data/nhanes/2017-2018_Demographics.XPT"))
```

### Subset Data:

Subset 'diet_behavior' as 'diet'

```{r}
diet <- select(diet_behavior, SEQN, DRD360, DRD370B, DRD370BQ, DRD370Q, DRD370QQ)
```

Subset 'diabetes' as 'tiid'

```{r}
tiid <- select(diabetes, SEQN, DIQ010)
```

Subset 'blood_hg' as 'bhg'

```{r}
bhg <- select(blood_hg, SEQN, LBXIHG, LBDIHGSI, LBXBGE, LBXBGM)
```

Subset "urine_hg' as 'uhg'

```{r}
uhg <- select(urine_hg, SEQN, URXUHG)
```

### Merge subset Data:

```{r}
df <- list(diet, tiid, bhg, uhg)

df <- df %>% reduce(full_join, by = 'SEQN')
```

1.  Filter Data Frame DF for the following:

```{r}
# Assuming your dataframe is named `nhanes_data`
df <- df %>%
  # Filter out rows where DIQ010 or DRD360 are NA
  filter(!is.na(DIQ010), !is.na(DRD370B)) %>%
  # Keep only rows where DIQ010 and DRD360 are 1 or 2
  filter(DIQ010 %in% c(1, 2), DRD370B %in% c(1, 2)) %>%
  # Recode 1 to "Yes" and 2 to "No" for DIQ010 and DRD360
  mutate(
    DIQ010 = ifelse(DIQ010 == 1, "Has Diabetes", "Naur Diabetes"),
    DRD370B = ifelse(DRD370B == 1, "Consumes Tuna? Yes", "Consumes Tuna? No")
  )
  
```

### Week **5**: **Notes about last weeks In-class Assignment:**

-   B \<- Has Diabetes

-   A \<- Eats Tuna

### P(B)

P(B) the probability that a person ***has diabetes*** (B) can be found by the following:

```{r}

df%>% 
  tabyl(DIQ010)

```

### L(B\|A) = P(A\|B) or (likelihood = probability)

P(A\|B) the probability that a person ***consumes tuna*** (A) given that a person ***has diabetes*** (B):

```{r}
df %>% 
  tabyl(DRD370B, DIQ010) %>% 
  adorn_percentages("col")
```

Reason as to why "col" and not "row"? We're looking for "diabetes" to be 100% given In other words we are looking to isolate a population given a reason, in this case the reason is that a person ***has*** diabetes, given P(A\|B) where B is equal to the term "has diabetes".

We can also do this and it means the same thing:

```{r}
df %>% 
  tabyl(DIQ010, DRD370B) %>% 
  adorn_percentages("row")
```

### P(A)

P(A) the probability that a person ***consumes tuna*** (A) can be found by the following:

```{r}
df %>% 
  tabyl(DRD370B)
```

### P(B\|A) 🎉 !

in other words, the question we have been trying to ask ourselves revealed that:

```{r}
probability_has_diabetes <- 0.13
probability_consumes_tuna_given_diabetes <- 0.36
probability_consumes_tuna <- 0.36
```

```{r}
(probability_has_diabetes * probability_consumes_tuna_given_diabetes) / probability_consumes_tuna

```
